Manuscript title here

### 2024-09-22 manuscript.Rmd
la till ett par källor

To differentiate the severity of trauma shock, we can divide shock into different degrees. One of such classification is the ATLS classification for major haemorrhage. The classification is divided into four classes, based on the patients: estimated bloodloss, heart frequency (HF), bloodpressure (BP), pulse pressure, respiratory rate (RR), mental status and urine output. However, this only meant to be used for hypovolemic shocks and not the other types. There is other classification for predicting massive transfusion (TASH, ABC), triaging (RETTS) or tracking the clinical deterioration of patients (NEWS2). The common denominator for these classifications is they are all roughly based on vital parameters, however none of them is specifically designed to classify trauma shock degree regardless of the type/cause of shock. ATLS in this regard, even though having its own shortcomings[@guly_vital_2011], still is one of the most used classifications for this topic(Källa).

@@

To qualify into the trauma registry, requires admission by trauma team activation or any admission with an Injury Severity Score (ISS) of more than nine. We included all patients in ... between 2014 and 2023  We excluded patients younger than 15 and/or were dead on arrival.

@@

<!-- Elaborate on the outcome OFI. How is it defined? What are the different types of OFI? -->

@@

<!-- Elaborate on how you will categorise the degree of shock, what are the different levels?-->
The first group will be patients presenting with a HR below 100 beats per minute(BPM), respiratory rate between 14 to 20. The original ATLS classification only states normal BP, but no number. In our case, we will be using the systolic over 90mmhg as the lower limit up to 65 years old, and 110mmhg as the lower limit for patients older (geriatric patients). The...

@@

All available data in the trauma care quality database? will be included, regarding the studied variables. <!-- Mention that you will include all available data-->

>>>>

To differentiate the severity of trauma shock, we can divide shock into different degrees. One of such classification is the ATLS classification for major haemorrhage. The classification is divided into four classes, based on the patients: estimated bloodloss, heart frequency (HF), bloodpressure (BP), pulse pressure, respiratory rate (RR), mental status and urine output. However, this only meant to be used for hypovolemic shocks and not the other types. There is other classifications for predicting massive transfusion (TASH,TBSS, ABC), triaging (RETTS) or tracking the clinical deterioration of patients (NEWS2). The common denominator for these classifications is they are all roughly based on vital parameters, however none of them is specifically designed to classify trauma shock degree regardless of the type/cause of shock. ATLS in this regard, even though having its own shortcomings[@guly_vital_2011], still is one of the most used classifications for this topic(Källa).

@@

To qualify into the trauma registry, requires admission by trauma team activation or any admission with an Injury Severity Score (ISS) of more than nine. We included all patients in "SWETRAU" and "Problem" between 2014 and 2023  We excluded patients younger than 15 and/or were dead on arrival.

@@

<!-- Elaborate on the outcome OFI. How is it defined? What are the different types of OFI? -->
The outcome was defined as presence of OFI, determined by the multidisciplinary M&M conference. An OFI can be expressed as various types of care errors, which include but not limited to communication, lack of resources, treatment delays, level of care, lack of competence and so on. In this study, two outcomes will be used, one is all of the types of OFI described, and the other is binary where OFI exists or does not exist.

@@

<!-- Elaborate on how you will categorise the degree of shock, what are the different levels?-->
Due to lack of data on FAST, type of injury and clinical chemistry, we will not be able to use categorisations such as TASH, TBSS and ABC with better sensitivity and specificity than ATLS[@yucel_trauma_2006][@ogura_modified_2016][@hanlin_prehospital_2024].

Version 1,
Due to the newer versions of ATLS classifications lack actual numbers, we will be using simplified and numerical version approximated by Dunham et al[@dunham_comparison_2017].

Version 2,
Due to the newer versions of ATLS classifications lack actual numbers, we will be approximating the values to numbers. For HF... For BP

The first group will be patients presenting with a HR below 100 beats per minute(BPM), respiratory rate between 14 to 20. The original ATLS classification only states normal BP, but no number. In our case, we will be using the systolic over 90mmhg as the lower limit up to 65 years old, and 110mmhg as the lower limit for patients older (geriatric patients). The...

Hur gör vi med de som hamnar mellan klasserna?

@@

<!-- Mention that you will include all available data-->
All available data in the trauma care quality database? will be included, regarding the studied variables.
